---
title: "Assignment1"
author: "George Paul"
date: "2024-01-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# install.packages('readxl')
library(readxl)

# install.packages("vioplot")
library("vioplot")

# install.packages("ggplot2")
library(ggplot2)
```

```{r}
rm(list = setdiff(ls(), lsf.str()))
```

```{r}
excel_path <- 'D:\\FILES\\BRSMa1.xlsx'
```

# Question 1

```{r}
q1data <- read_excel(excel_path)
q1data
```

```{r}
boxplot(q1data)
```

```{r}
vioplot(q1data)
```

```{r}
plot(q1data$x1, type = "l") +
  lines(q1data$x2, type = "l", col = "red") +
  lines(q1data$x3, type = "l", col = "green") +
  lines(q1data$x4, type = "l", col = "blue")

```

The box plot for this data is entirely misleading - showing the same plots for all data groups.

The violin plot does away with most problems giving a much better idea of the distributions but perhaps x1 and x3 seem too similar even though they are very different.

The line plot is also a valid method of showing the data since the data seems to be of a somewhat continuous nature and does away with the problems of the violin plot.

# Question 2

```{r}
q2data <- read_excel(excel_path, sheet = "Movement Personality Results")
q2data
```

```{r}
mosaicplot(as.data.frame(t(q2data)), main = "Personality and Motion", color = "plum", las = 2, cex.axis = 0.5)
```

```{r}
plot(q2data$Openness, type = "o",col = "blue", xlab = "Joints", ylab = "Jointimportance values", main = "Personality and Motion") + 
lines(q2data$Conscientiousness, col="red") + 
lines(q2data$Extraversion, col="green") + 
lines(q2data$Agreeableness, col="yellow") +
lines(q2data$Neuroticism, col="purple")

```

Since we want to know how much each joint's value contributes to that personality trait, a mosaic plot with tiles scaled according to contribution is a great choice

An inferior way to present this data is shown above as a line plot. The data is not continuous nor does it show the relation each point will have with their counterparts in the same joint or personality trait.

# Question 3.1

```{r}
q31labels <- c("Zombie Males", "Safe Males", "Zombie Females", "Safe Females")
q31colors <- c("steelblue4", "steelblue1", "violetred4", "violetred1")
q31.1 <- c(118, 62, 4,141)
pie(q31.1, labels = q31labels, col = q31colors, main="Safe Zone")
```

```{r}
q31.2 <- c(154, 25, 13, 93)
pie(q31.2, labels = q31labels, col = q31colors, main="Contaminated City")
```

```{r}
q31.3 <- c(422, 88, 106, 90)
pie(q31.3, labels = q31labels, col = q31colors, main="Rural Area")
```

```{r}
q31.4 <- c(670, 192, 3, 20)
pie(q31.4, labels = q31labels, col = q31colors, main="Isolated Island")
```

Pie charts can easily show you the fraction of people that survived in each zone and importantly, also the relation between being male or female in that zone.

# Question 3.2

```{r}
q32data <- read_excel(excel_path, sheet = 3)[c("RI", "Type")]
boxplot(q32data$RI ~ q32data$Type)
```
