---
title: "BRSM Results Analysis"
author: "George Paul"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = setdiff(ls(), lsf.str()))
```

```{r}
# install.packages('readxl')
library(readxl)
```

```{r}
excel_path <- 'D:\\FILES\\BRSM_Results Visualization.xlsx'
data <- read_excel(excel_path)
data
nums <- data[["No. of Mosquitoes"]]
nums
grps <- data[["Group"]]
grps
count_b <- sum(grps == "Beer")
count_b
count_w <- sum(grps == "Water")
count_w
beer_list <- subset(data, Group == "Beer")
beer_list
watr_list <- subset(data, Group == "Water")
obs_stat_med <- median(beer_list$`No. of Mosquitoes`) - median(watr_list$`No. of Mosquitoes`)
obs_stat_med
obs_stat_t <- t.test(watr_list$`No. of Mosquitoes`, beer_list$`No. of Mosquitoes`)
as.double(obs_stat_t$statistic)
```

## Question 2.a

```{r}
get_group_median_diff <- function() {
  beer_sample <- sample(nums, count_b, replace = TRUE)
  watr_sample <- sample(nums, count_w, replace =TRUE)
  median(beer_sample) - median(watr_sample)
}
get_group_median_diff()
```

```{r}
counter <- 0
iter_count <-10000 

medians <- c()

repeat {
  medians <- c(medians, get_group_median_diff())

  counter <- counter + 1
  if (counter >= iter_count) {
    break
  }
}

plot(hist(medians, main = "Median of Beer sample - Median of Water sample"))+
abline(v = obs_stat_med, col = "red", lty = 2)
```

```{r}
p_val_qa <- sum(medians >= obs_stat_med) / length(medians)
p_val_qa
```

The calculated p-value is: $\frac{\text{Number of values where val }\ge\text{ Observed Statistic}}{\text{Total Number of values}}$

This was calculated to be $\text{p-value} \approx 0.01 < \alpha$ . Hence we can conclude that it is statistically significant.

#### 2.c for (a) step statistic

```{r}
p_val_qa_nondir <- (sum(medians >= obs_stat_med) + sum(medians <= -obs_stat_med)) / length(medians)
p_val_qa_nondir
```

The calculated p-value for non-directional hypothesis is: $\frac{\text{Number of values where val }\ge\text{ Observed Statistic} + \text{Number of values where val }\le(-\text{ Observed Statistic})}{\text{Total Number of values}}$

This was calculated to be $\text{p-value} \approx 0.02 < \alpha$ . Hence we can conclude that it is statistically significant.

## Question 2.b

```{r}
get_group_t <- function() {
  beer_sample <- sample(nums, count_b, replace = TRUE)
  watr_sample <- sample(nums, count_w, replace =TRUE)
  as.double(t.test(beer_sample, watr_sample)$statistic)
}
get_group_t()
```

```{r}
counter <- 0
iter_count <-10000 

ts <- c()

repeat {
  ts <- c(ts, get_group_t())

  counter <- counter + 1
  if (counter >= iter_count) {
    break
  }
}

plot(hist(ts, main = "T statistic"))
# abline(v = obs_stat_t, col = "red", lty = 2)
```

## Question 3

```{r}
excel_path <- 'D:\\FILES\\iqdata.xlsx'
data <- read_excel(excel_path)
data
```

```{r}
sco_list <- data$`Placement \r\nTESTSCORE`
iqs_list <- data$IQ
original_corr <- cor(sco_list, iqs_list)
original_corr
```

```{r}
counter <- 0
iter_count <-10000 

corrs <- c()

repeat {
  shuff_sco_list <- sco_list[sample(length(sco_list))]
  shuff_iqs_list <- iqs_list[sample(length(iqs_list))]
  cor(shuff_sco_list, shuff_iqs_list)
  corrs <- c(corrs, cor(shuff_sco_list, shuff_iqs_list))

  counter <- counter + 1
  if (counter >= iter_count) {
    break
  }
}

plot(hist(corrs, main = "Correlation bootstrap distribution"))+
abline(v = original_corr, col = "red", lty = 2)
```

As it stands in the latest simulation, the originally calculated correlation is beyond the range of any of the randomly generated data. Hence the calculated p-value will be $\approx 0 \le \alpha = 0.05$ . We can conclude that the correlation that was found was statistically quite significant.
